# Matricies

Matricies are the most common data format bioinformaticians work with (microarray/RNAseq data for example). Lets make one:

```{r}
m <- matrix(0,ncol=5,nrow=10)
m
```
This will create a matrix filled with zeros. To transpose (flip) the matrix we use `t()` (this will be important later!)

```{r}
tposed.m <- t(m)
tposed.m
```

## Subsetting

Lets make a matrix (and a vector) containing integer values so we can take a look at how subsetting work in R:
```{r}
v <- 1:10
m <- t(matrix(1:50,ncol=10,nrow=5))
m
```
We can access individual elements using square brackets `[]`. Here are some examples:

```{r}
v[c(7,1,5)] #access elements 7 1 and 5 of the vector
m[1,] # access the first row of the matrix
m[,3] # the 3rd column
m[8,2] # the value in the 8th row and 2nd column
m[3:7,4] # the 3rd to 7th row of the 4th column
```

Note the `c(7,1,5)` where we subset vector `v`. `c` means *combine* and it allows element in the `()` to be collected into a single vector.

We often have the need to collect vectors and assemble them into a matrix. This can be done using the `rbind` (row) and `cbind` (column) function:

```{r}
v1 <- 1:10
v2 <- 101:110
rbound.mat <-rbind(v1,v2)
cbound.mat <- cbind(v1,v2)
rbound.mat
cbound.mat
```
