# Functions

Functions and chunks of code that execture several lines of code to perform a task. Once you have a few lines of useful code that you want to apply repeatedly, a function is a nice way to wrap them up so it can be used quickly when needed. lets turn the code you wrote in the exercise earlier into a function where we also calculate the variance for a gene too.

```{r}
calc.mean.and.sd <- function(mat){
  
  mn <- NULL
  vr <- NULL
  
  for(i in 1:nrow(mat)){  
      mn <- c(mn,mean(mat[i,]))
      vr <- c(vr,var(mat[i,]))
    
  }
  res <- list(mns=mn,vars=vr)
  res # the last line in a function is what the function will return
}
```

You can see a loop is started, and the output from each loop is put into variables `mn` and `vr`. These are then put into a list which is was it returned at the end.
By putting this code into a function we can now calculate the means and deviations of any matrix. For example, `gmp`:

```{r}
gmp.mn.sd <- calc.mean.and.sd(gmp) 
```

Functions can also work with built in conditions:

```{r}
animal.maths <- function(value1,value2,animal=c("pig","cow")){
  
  if(animal=="pig"){print(value1/value2)}
  if(animal=="cow"){print(value1*value2)}
  
}
animal.maths(5,5,"pig")
animal.maths(5,5,"cow")

```

The above can be simplified a bit further since there are only two options `pig` and `cow` by using the `else` statement:

```{r}
animal.maths <- function(value1,value2,animal=c("pig","cow")){
  
  if(animal=="pig"){print(value1/value2)}
  else{print(value1*value2)}
  
}
animal.maths(5,5,"pig")
animal.maths(5,5,"cow")

```

In the case above we're only concerned if one of the arguments is `pig`, anything else goes:

```{r}
animal.maths(5,5,"dog")
```


These functions can now be "banked" for use whenever they are needed (probably not `animal.maths` to be fair). *However*, you should avoid using for-loops etc altogether since R has some built in functions that are much quicker and tidier. Lets look at that now.
