# Apply
'apply' is a commonly used function in R to speed up matrix calculation. For example, to calculates means of a matrix we can do this:
```{r}
lthsc.row.mn <- apply(lthsc, 1, mean) # means of rows
lthsc.col.mn <- apply(lthsc, 2, mean) # means of columns
```
The format for the function is therefore (1) the matrix, (2) the direction in which you would like to apply a function, and (3) the function to be applied.

***Exercise:*** Use `apply` to calculate row and column totals and deviations for a yeast dataset of your choosing.

Your own functions can also be used with `apply` when used as the 3rd argument. Example:
```{r}

example.func <- function(v){
  
  val <- (mean(v)*sd(v))/sum(v) ## This is a nonsense operation.
  val
}

ex.apply <- apply(mep, 1, example.func)

```

Lets use the apply function to get the top 500 most variable genes in our HSPC dataset:
```{r}
gene.vars <- apply(hspc.data, 1, var)
top.var.genes <- names(rev(sort(gene.vars))[1:500])
hspc.var <- hspc.data[top.var.genes,]
```

