# Kmeans

We're going to program our first algorithm together! Kmeans ckustering is a common way to cluster expression data, and the algorithm is actually pretty simple, and a great start to learning programmatic thinking. The data we'll use comes from yeast where the expression of 256 genes have been measured as a synchronised population go through two divisions. It's a nice, small dataset with clear clusters.

There is a file called ""Spellman_Yeast_Cell_Cycle.tsv". Load it into a variable called `ycc` and convert it to a matrix.

```{r}
ycc <- read.delim("Spellman_Yeast_Cell_Cycle.tsv",row.names=1,header=T,sep="\t")
ycc <- as.matrix(ycc)
```

The data has already been z-scored, so we need to learn how to calculate the Euclidean distance between 2 genes (vector). We do this using the formula:

$$\Sigma{(G_1-G_2)^2}$$

Write a function called `e.dist` which takes two vectors `v1` and `v2` and calculates the distance between them.

```{r,class.source = 'fold-hide',results=F}
e.dist <- function(v1,v2){
  d <- sum((v1-v2)^2)
  d
}
```

The kmeans algorithm needs the user to say how many clusters are being searched for, so in this case we'll say **10**.

1. Define 10 centers randomly, i.e pick 10 genes at random.
2. Calculate the distance between each gene to each of the 10 centroids.
3. Assign each gene to the closest centroid.
4. For each of the 10 clusters, calculate a new centroid.
5. Repeat steps 2,3, and 4 100 times.

We'll now work through this problem together and crowd source a solution.

```{r}
centroids <- ycc[sample(1:256,10,replace=F),]

dists <- matrix(0,nrow=256,ncol=10)

clusters <- NULL

for(rounds in 1:100){

  for(i in 1:256){
  
    for(j in 1:10){
      dists[i,j] <- e.dist(ycc[i,],centroids[j,])
    }
    
  }

  clusters <- apply(dists,1,which.min)

  for(k in 1:10){
    centroids[k,] <- apply(ycc[which(clusters==k),],2,mean)
  }
}
```

***Exercise*** Plot your clusters using base R or ggplots. Dealers choice.

```{r,class.source = 'fold-hide',results=F}
library(reshape2)
ycc.cl <- cbind(as.factor(clusters),ycc)
colnames(ycc.cl)[1] <- c("Cluster")

ycc.df <- data.frame(ID=rownames(ycc.cl), ycc.cl, row.names=NULL)

ytm <- melt(ycc.df,c("ID","Cluster"))

ytm <- melt(as.data.frame(ycc.cl))

ytm[1,]

ggplot(ytm,aes(x = variable, y = value,group = ID)) + geom_line()+facet_wrap(~ Cluster, ncol = 3)


```


***Improvements***

1. The code above will do 100 iterations, but what if the algorithm converges after 20? Alter the code so it stops when no further changes are made and help save the planet.
2. Write a function from this code that will take any `data`, number of clusters `K`, and `iterations` as an argument and perform Kmeans clustering. Return a list that contains the cluster assignments and data.
3. Write a function that will plot from the list, a cluster of the user's choice.


